Re-examining cross-cultural similarity judgments using lexical co-occurrence

### 2023-02-02 code/4_paper/manuscript/manuscript.Rmd
manuscript updates

---
title: "Re-examining cross-cultural similarity judgments using lexical co-occurrence"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1, T2A, T5]{fontenc}
  - \usepackage{kotex}
author-information: > 
    \author{{\large \bf Khuyen N. Le (knl005@ucsd.edu)}$^{1}$, {\large \bf Alexandra Carstensen (abc@ucsd.edu)}$^{1}$, \\ {\large \bf Shan Gao (shangaocog@gmail.com)$^2$}, {\large \bf Michael C. Frank (mcfrank@stanford.edu)$^3$},  \AND
    $^1$Department of Psychology, University of California, San Diego, $^2$Department of Psychology, University of Chicago, \\ $^3$Department of Psychology, Stanford University
    }
abstract: >
 Is “cow” more closely related to “grass” or “chicken”? Speakers of different languages judge similarity in this context differently, but why? One possibility is that cultures co-varying with these languages induce differences in conceptualizations of similarity. Specifically, East Asian cultures may promote reasoning about thematic similarity, by which cow and grass are more related, whereas Western cultures may bias judgments toward taxonomic relations, like cow-chicken. This difference in notions of similarity is the consensus interpretation for cross-cultural variation in this paradigm. We consider, and provide evidence for, an alternative possibility, by which notions of similarity are similar across contexts, but the statistics of the environment vary. On this account, similarity judgments are guided by co-occurrence in experience, and hearing about cows and grass or cows and chickens more often could induce preferences for these groupings, and account in part for apparent differences in notions of similarity across contexts.

@@

```{r setup, include = FALSE, cache=F}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")
library("lme4")
library("emmeans")
library("patchwork")
library("effsize")
library("kableExtra")
library("irr")
require(agreement)
library("report")
library("MuMIn")

@@

An influential perspective within cultural psychology links these differences in similarity judgment to tendencies toward analytic processing in Western cultures, and holistic processing in East Asian cultures [@Nisbett2003].
Analytic processing emphasizes rule-like relationships predicated on objects and their properties and, correspondingly, taxonomic similarity, while holistic processing emphasizes relations between objects and their context and therefor, thematic similarity.
In related work, East Asian participants show a higher level of sensitivity to context than their Western counterparts when reproducing drawings from memory [@Ji2000]; visually exploring naturalistic scenes [@Chua2005]; describing scenes [@Masuda2001]; and in explaining the causes of ambiguous behaviors [@Choi1999]. The consensus interpretation in this literature ascribes cross-cultural differences in similarity judgment to variation in the *conceptualization* of similarity -- with people from East Asian cultures relying on a more thematic notion of similarity than Westerners.

@@

Our study uses cosine distances of fastText word vectors as a measure of lexical co-occurrence[^2].
fastText is a system that uses lexical co-occurrence information to generate a vector representing each word in its lexicon [@Mikolov2018].
fastText has also been shown to be sensitive to cultural differences in word meanings: @Thompson2020 demonstrated that the distribution of semantic meaning clusters generated by fastText trained on language-specific corpora correlates with the cultural, historical, and geographical similarities of these languages.
<!-- This proof of concept provides support for our use of fastText in comparing different languages with varying levels of relatedness.-->

[^2]: We also carried out our analysis using raw lexical co-occurrences and obtained similar results.

We note that language statistics may incorporate both cultural-specific environmental statistics, that is, the experiential *input* to similarity judgments, and culture-specific senses of similarity, the *conceptualization* of similarity.
For example, cultural features (like farming) can lead to differences in environmental statistics (seeing cows and grass) and these can influence language (talking more about cows and grass).
But other cultural features (like conceiving of similarity thematically) could also cause individuals to talk differently about the same experiences (mentioning what cows eat rather than what other animals cows are like).
Acoordingly, our approach examines the extent to which language statistics can predict cross-cultural differences in similarity judgments with the understanding that language statistics are likely a proxy for both *input* to and *conceptualization* of similarity.

@@

```{=html}
<!--@Ji2004 suggest that the notion of similarity is what varies across cultural groups, but it is possible that similarity is regarded in the same way and it is actually the input to this similarity judgment that varies across cultural contexts.
It may be possible to gain traction on potential mechanisms by examining whether variation in similarity judgments co-vary with environmental statistics that differ across cultural and linguistic contexts.--
```
The present study tests whether the statistics of language can be used to predict cross-cultural differences in similarity judgments, and particularly, whether this approach provides additional insight beyond cross-cultural characterizations based on taxonomic vs thematic preferences.

We measured taxonomic versus thematic similarity matching in a forced-choice word triad task in three populations. Following @Ji2004, we measured preferences in the US and China. In addition, we collected data in a novel context: Vietnam. Vietnam is a Southeast Asian country that borders China and has historically been greatly influenced by Chinese culture [@Hui2002].
Therefore, it serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies -- that Eastern and Western cultures have different notions of similarity -- extends beyond mainland China. In addition to these replication and extension questions, we tested whether fastText vectors from corpora corresponding to each language context (English, Mandarin, and Vietnamese) are good predictors for similarity judgments in each population.

@@

Data collection and analyses for this study were pre-registered, and the pre-registration is available at [blinded for peer review]. We recruited `r df.excl_US$ppts_finished` participants from the US, `r df.excl_VN$ppts_finished` from Vietnam, and `r df.excl_CN$ppts_finished` from mainland China.
All participants were recruited through snowball sampling, which, in the US was seeded with student email lists at a large university, in Vietnam (VN) with Vietnam-based student groups on Facebook, and in China (CN) with student groups on WeChat.
US participants were compensated with \$5 gift certificates (USD), VN participants received 50,000₫ (VND) in phone credit, and CN participants received 25CNY through WeChat credit transfer.

We excluded `r df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl` US participants, `r df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl` VN participants, and `r df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl` CN participants who missed two or more of the 10 attention check questions.

In doing so, we deviated from our preregistered exclusion criterion, which used a more stringent policy, excluding participants who missed any of the 10 attention checks. This approach would have substantially reduced our sample size, disproportionately affecting the VN sample (resulting in a final sample of 109 US, 132 CN, and 57 VN participants).
We report results with the less stringent exclusion criterion, but we performed analyses with both samples and found broadly similar results. Any differences are noted in the Analysis.

We also applied four demographic exclusion criteria designed to reduce cross-cultural influences across our samples: we excluded participants who (1) were not native speakers of the test language (English, Vietnamese, or Mandarin Chinese), (2) were fluent in another one of the study languages, (3) had lived outside of the test country for more than two years, or (4) had significant international experience (more than 6 international experiences of 2 days or longer). Following our pre-registration, we dropped any exclusion criterion that would exclude 25% or more of any one population, but only for that population. Specifically, we did not exclude VN and CN participants who spoke English, or US participants with significant international experience.
As a result of these demographic exclusions, we excluded `r df.excl_US$ppts_after_att_check_excl - df.excl_US$ppts_after_demog_excl` US, `r df.excl_VN$ppts_after_att_check_excl - df.excl_VN$ppts_after_demog_excl` VN, and `r df.excl_CN$ppts_after_att_check_excl - df.excl_CN$ppts_after_demog_excl` CN participants.
The final US sample included `r df.excl_US$ppts_after_demog_excl` participants (`r df.gender_E %>% filter(gender == "Male") %>% pull(n)`M, `r df.gender_E %>% filter(gender == "Female") %>% pull(n)`F, `r df.gender_E %>% filter(gender == "Non-binary") %>% pull(n)` non-binary, `r df.gender_E %>% filter(gender == "Decline to answer") %>% pull(n)` other), with mean age = `r round(df.age_E$mean_age, 2)` (SD = `r round(df.age_E$sd_age, 2)`).
Our VN sample included `r df.excl_VN$ppts_after_demog_excl` participants (`r df.gender_V %>% filter(gender == "Nam") %>% pull(n)`M, `r df.gender_V %>% filter(gender == "Nữ") %>% pull(n)`F, `r df.gender_V %>% filter(gender == "Từ chối trả lời") %>% pull(n)` other), with mean age = `r round(df.age_V$mean_age, 2)` (SD = `r round(df.age_V$sd_age, 2)`).
And the CN sample included `r df.excl_CN$ppts_after_demog_excl` participants (`r df.gender_M %>% filter(gender == "男性") %>% pull(n)`M, `r df.gender_M %>% filter(gender == "女性") %>% pull(n)`F, `r df.gender_M %>% filter(gender == "拒绝回答") %>% pull(n)` other), with mean age = `r round(df.age_M$mean_age, 2)` (SD = `r round(df.age_M$sd_age, 2)`).[^4]

[^4]: A table summarizing the number of participants omitted for each criterion is included in the Supplementary Information.

@@

All stimuli and scripts for the experiment and analysis are available at [BLINDED FOR REVIEW]. We collected and adapted stimuli from previous studies to create test triads consisting of a cue with one thematic and one taxonomic match option.
For example, "cow," "grass," and "chicken," where "cow" is the cue, "grass" is the thematic match, and "chicken" the taxonomic match.
We included 105 such triads, a superset including triads pulled from supplemental information and in-text examples across the literature, and others that we adapted or created <!--(see SI for the full list of stimuli and sources)-->.
We selected triads on the basis of cultural familiarity within the US, Vietnam, and China.
The triads were translated from English to Vietnamese and Mandarin by fluent bilingual speakers of each language, and backtranslated to English by another fluent bilingual who was naive to the original English version. The backtranslations were checked against the original English to identify ambiguities in the translation, which were resolved through discussion and selection of alternative terms when relevant.

## Procedure

Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue ("Which thing is most closely related to the bolded item?" and translated equivalents).
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 10 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 10 attention check questions, with all items presented in randomized orders that varied between subjects.

@@

In keeping with @Ji2004, we would expect participants from mainland China to prefer thematic matches more than US participants (e.g., to prefer the cow-grass match over cow-chicken to a greater extent than US participants).
On the basis of the previous literate more broadly, we would also expect participants from Vietnam to pattern with China, showing a stronger preference for thematic matches than US participants.

We observed the strongest preference for thematic matching in China (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), and then the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`; see Figure 1).

@@

To test for differences in similarity judgment between the countries, we ran a mixed-effects logistic regression predicting triad match (taxonomic or thematic) with country (US, China, or Vietnam) as a fixed effect.
As random effects, we included an intercept per subject and one per triad, as well as by-triad random slopes for country to account for variation in the country effect across triads.

We found a significant effect of country ($\chi^2$(`r fit.country.anova["country", "Df"]`) = `r round(fit.country.anova["country", "Chisq"], 2)`, p `r ifelse( fit.country.anova["country", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.country.anova["country", "Pr(>Chisq)"], 3)), "< .001")`), but this effect is driven by the difference between US and CN responding ($\beta$ = `r round(fixef(fit.country)["countryUS"], 2)`, p `r ifelse(as.data.frame(summary(fit.country)[["coefficients"]])["countryUS", "Pr(>|z|)"] > 0.001, paste0("= ", round(as.data.frame(summary(fit.country)[["coefficients"]])["countryUS", "Pr(>|z|)"], 3)), "< .001")`). There was no statistical difference between matching in Vietnam and China ($\beta$ = `r round(fixef(fit.country)["countryVietnam"], 2)`, p `r ifelse(as.data.frame(summary(fit.country)[["coefficients"]])["countryVietnam", "Pr(>|z|)"] > 0.001, paste0("= ", round(as.data.frame(summary(fit.country)[["coefficients"]])["countryVietnam", "Pr(>|z|)"], 3)), "< .001")`), or the US and Vietnam ($\beta$ = `r round(fixef(fit.country_EN_VN)["countryVietnam"], 2)`, p `r ifelse(as.data.frame(summary(fit.country_EN_VN)[["coefficients"]])["countryVietnam", "Pr(>|z|)"] > 0.001, paste0("= ", round(as.data.frame(summary(fit.country_EN_VN)[["coefficients"]])["countryVietnam", "Pr(>|z|)"], 3)), "< .001")`).

@@

For US responding: only the English (EN) corpus is a significant predictor[^5].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^5]: With preregistered exclusion criterion: only the English (EN) and Mandarin (ZH) corpus are significant predictors.

For Vietnam responding: only the Vietnamese (VI) and Mandarin (ZH) corpus are significant predictors[^6].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

>>>>

---
title: "Re-examining cross-cultural similarity judgments using lexical co-occurrence"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1, T2A, T5]{fontenc}
  - \usepackage{kotex}
  - \usepackage{CJKutf8}
author-information: > 
    \author{{\large \bf Khuyen N. Le (knl005@ucsd.edu)}$^{1}$, {\large \bf Alexandra Carstensen (abc@ucsd.edu)}$^{1}$, \\ {\large \bf Shan Gao (shangaocog@gmail.com)$^2$}, {\large \bf Michael C. Frank (mcfrank@stanford.edu)$^3$},  \AND
    $^1$Department of Psychology, University of California, San Diego, $^2$Department of Psychology, University of Chicago, \\ $^3$Department of Psychology, Stanford University
    }
abstract: >
 Is “cow” more closely related to “grass” or “chicken”? Speakers of different languages judge similarity in this context differently, but why? One possibility is that cultures co-varying with these languages induce differences in conceptualizations of similarity. Specifically, East Asian cultures may promote reasoning about thematic similarity, by which cow and grass are more related, whereas Western cultures may bias judgments toward taxonomic relations, like cow-chicken. This difference in notions of similarity is the consensus interpretation for cross-cultural variation in this paradigm. We consider, and provide evidence for, an alternative possibility, by which notions of similarity are similar across contexts, but the statistics of the environment vary. On this account, similarity judgments are guided by co-occurrence in experience, and hearing about cows and grass or cows and chickens more often could induce preferences for these groupings, and account in part for apparent differences in notions of similarity across contexts.

@@

```{r setup, include = FALSE, cache=F}
library("papaja")
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("car")            
library("emmeans")
library("patchwork")
library("effsize")
library("kableExtra")
library("irr")
require(agreement)
library("report")
library("MuMIn")
library("lme4")

@@

An influential perspective within cultural psychology links these differences in similarity judgment to tendencies toward analytic processing in Western cultures, and holistic processing in East Asian cultures [@Nisbett2003].
Analytic processing emphasizes rule-like relationships predicated on objects and their properties and, correspondingly, taxonomic similarity, while holistic processing emphasizes relations between objects and their context and therefore, thematic similarity.
In related work, East Asian participants show a higher level of sensitivity to context than their Western counterparts when reproducing drawings from memory [@Ji2000]; visually exploring naturalistic scenes [@Chua2005]; describing scenes [@Masuda2001]; and in explaining the causes of ambiguous behaviors [@Choi1999].
The consensus interpretation in this literature ascribes cross-cultural differences in similarity judgment to variation in the *conceptualization* of similarity -- with people from East Asian cultures relying on a more thematic notion of similarity than Westerners.

@@

Our study uses cosine distances of fastText word vectors as a measure of lexical co-occurrence[^1].
fastText is a system that uses lexical co-occurrence information to generate a vector representing each word in its lexicon [@Mikolov2018].
fastText has also been shown to be sensitive to cultural differences in word meanings: @Thompson2020 demonstrated that the distribution of semantic meaning clusters generated by fastText trained on language-specific corpora correlates with the cultural, historical, and geographical similarities of these languages.
<!-- This proof of concept provides support for our use of fastText in comparing different languages with varying levels of relatedness.-->

[^1]: We also carried out our analysis using raw lexical co-occurrences and obtained similar results.

We note that language statistics may incorporate both cultural-specific environmental statistics, that is, the experiential *input* to similarity judgments, and culture-specific senses of similarity, the *conceptualization* of similarity.
For example, cultural features (like farming) can lead to differences in environmental statistics (seeing cows and grass) and these can influence language (talking more about cows and grass).
But other cultural features (like conceiving of similarity thematically) could also cause individuals to talk differently about the same experiences (mentioning what cows eat rather than what other animals cows are like).
Accordingly, our approach examines the extent to which language statistics can predict cross-cultural differences in similarity judgments with the understanding that language statistics are likely a proxy for both *input* to and *conceptualization* of similarity.

@@

```{=html}
<!--@Ji2004 suggest that the notion of similarity is what varies across cultural groups, but it is possible that similarity is regarded in the same way and it is actually the input to this similarity judgment that varies across cultural contexts.
It may be possible to gain traction on potential mechanisms by examining whether variation in similarity judgments co-vary with environmental statistics that differ across cultural and linguistic contexts.--
```
The present study tests whether the statistics of language can be used to predict cross-cultural differences in similarity judgments, and particularly, whether this approach provides additional insight beyond cross-cultural characterizations based on taxonomic vs. thematic preferences.

We measured taxonomic vs. thematic similarity matching in a forced-choice word triad task in three populations.
Following @Ji2004, we measured preferences in the US and China.
In addition, we collected data in a novel context: Vietnam.
Vietnam is a Southeast Asian country that borders China and has historically been greatly influenced by Chinese culture [@Hui2002].
Therefore, it serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies -- that Eastern and Western cultures have different notions of similarity -- extends beyond mainland China.
In addition to these replication and extension questions, we tested whether fastText vectors from corpora corresponding to each language context (English, Mandarin, and Vietnamese) are good predictors for similarity judgments in each population.

@@

Data collection and analyses for this study were pre-registered, and the pre-registration is available at [blinded for peer review].
We recruited `r df.excl_US$ppts_finished` participants from the US, `r df.excl_CN$ppts_finished` from mainland China, and `r df.excl_VN$ppts_finished` from Vietnam.
All participants were recruited through snowball sampling, which, in the US was seeded with student email lists at a large university, in China (CN) with student groups on WeChat, and in Vietnam (VN) with Vietnam-based student groups on Facebook.
US participants were compensated with \$5 gift certificates (USD), CN participants received 25CNY through WeChat credit transfer, and VN participants received 50,000₫ (VND) in phone credit,

We excluded `r df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl` US participants, `r df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl` CN participants, and `r df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl` VN participants, who missed two or more of the 11 attention check questions.

In doing so, we deviated from our preregistered exclusion criterion, which used a more stringent policy, excluding participants who missed any of the 10 attention checks.
This approach would have substantially reduced our sample size, disproportionately affecting the VN sample (resulting in a final sample of 109 US, 132 CN, and 57 VN participants).
We report results with the less stringent exclusion criterion, but we performed analyses with both samples and found broadly similar results.
Any differences are noted in the Results & Discussion section.

We also applied four demographic exclusion criteria designed to reduce cross-cultural influences across our samples: we excluded participants who (1) were not native speakers of the test language (English, Vietnamese, or Mandarin Chinese), (2) were fluent in another one of the study languages, (3) had lived outside of the test country for more than two years, or (4) had significant international experience (more than 6 international experiences of 2 days or longer).
Following our pre-registration, we dropped any exclusion criterion that would exclude 25% or more of any one population, but only for that population.
Specifically, we did not exclude VN and CN participants who spoke English, or US participants with significant international experience.
As a result of these demographic exclusions, we excluded `r df.excl_US$ppts_after_att_check_excl - df.excl_US$ppts_after_demog_excl` US, `r df.excl_CN$ppts_after_att_check_excl - df.excl_CN$ppts_after_demog_excl` CN, and `r df.excl_VN$ppts_after_att_check_excl - df.excl_VN$ppts_after_demog_excl` VN participants.
The final US sample included `r df.excl_US$ppts_after_demog_excl` participants (`r df.gender_E %>% filter(gender == "Male") %>% pull(n)`M, `r df.gender_E %>% filter(gender == "Female") %>% pull(n)`F, `r df.gender_E %>% filter(gender == "Non-binary") %>% pull(n)` non-binary, `r df.gender_E %>% filter(gender == "Decline to answer") %>% pull(n)` other), with mean age = `r round(df.age_E$mean_age, 2)` (SD = `r round(df.age_E$sd_age, 2)`).
The CN sample included `r df.excl_CN$ppts_after_demog_excl` participants (`r df.gender_M %>% filter(gender == "男性") %>% pull(n)`M, `r df.gender_M %>% filter(gender == "女性") %>% pull(n)`F, `r df.gender_M %>% filter(gender == "拒绝回答") %>% pull(n)` other), with mean age = `r round(df.age_M$mean_age, 2)` (SD = `r round(df.age_M$sd_age, 2)`).
And the VN sample included `r df.excl_VN$ppts_after_demog_excl` participants (`r df.gender_V %>% filter(gender == "Nam") %>% pull(n)`M, `r df.gender_V %>% filter(gender == "Nữ") %>% pull(n)`F, `r df.gender_V %>% filter(gender == "Từ chối trả lời") %>% pull(n)` other), with mean age = `r round(df.age_V$mean_age, 2)` (SD = `r round(df.age_V$sd_age, 2)`).[^2]

[^2]: A table summarizing the number of participants omitted for each criterion is included in the Supplementary Information.

@@

All stimuli and scripts for the experiment and analysis are available at [BLINDED FOR REVIEW].
We collected and adapted stimuli from previous studies to create test triads consisting of a cue with one thematic and one taxonomic match option.
For example, "cow," "grass," and "chicken," where "cow" is the cue, "grass" is the thematic match, and "chicken" the taxonomic match.
We included 105 such triads, a superset including triads pulled from supplemental information and in-text examples across the literature, and others that we adapted or created <!--(see SI for the full list of stimuli and sources)-->.
We selected triads on the basis of cultural familiarity within the US, China, and Vietnam.
The triads were translated from English to Mandarin and Vietnamese by fluent bilingual speakers of each language, and backtranslated to English by another fluent bilingual who was naive to the original English version.
The backtranslations were checked against the original English to identify ambiguities in the translation, which were resolved through discussion and selection of alternative terms when relevant.

## Procedure
\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue (following Ji et al., 2004): "Which thing is most closely related to the bolded item?" / "哪个事物与加粗项关联最紧密?" / "Thứ nào liên quan nhất với thứ được in đậm?".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 11 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 11 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}

@@

Following previous work, we would expect participants from mainland China to prefer thematic matches more than US participants (to use the cow-grass-chicken triad: we would expect participants from mainland China to prefer the cow-grass match over the cow-chicken match to a larger extent compared to US participants).
We would also expect that participants from Vietnam would pattern with China, and also show a significantly higher preference for thematic matches compare to US participants.

The group means of proportion of thematic response in mainland China is the highest (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by the groups means in Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), which is slightly higher than that of the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`) (Figure 1).

@@

To test for cross-context differences in similarity judgments between the countries, we ran a mixed-effects logistic regression predicting triad responding (taxonomic or thematic) with country (US, China, or Vietnam) as a fixed effect.
As random effects, we included an intercept per subject and one per triad, as well as by-triad random slopes for country to account for variation in the country effect across triads.

Overall, there is a significant effect of country on proportion of thematic responses ($\chi^2$(`r fit.country.anova["country", "Df"]`) = `r round(fit.country.anova["country", "Chisq"], 2)`, p `r ifelse( fit.country.anova["country", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.country.anova["country", "Pr(>Chisq)"], 3)), "< .001")`). However, this effect is driven by the difference between US and China responding ($\beta_{US}$ = `r round(fixef(fit.country)["countryUS"], 2)`, p `r ifelse(as.data.frame(summary(fit.country)[["coefficients"]])["countryUS", "Pr(>|z|)"] > 0.001, paste0("= ", round(as.data.frame(summary(fit.country)[["coefficients"]])["countryUS", "Pr(>|z|)"], 3)), "< .001")`). There is no statistical difference between the Vietnam and China responding ($\beta_{Vietnam}$ = `r round(fixef(fit.country)["countryVietnam"], 2)`, p `r ifelse(as.data.frame(summary(fit.country)[["coefficients"]])["countryVietnam", "Pr(>|z|)"] > 0.001, paste0("= ", round(as.data.frame(summary(fit.country)[["coefficients"]])["countryVietnam", "Pr(>|z|)"], 3)), "< .001")`), and the US and Vietnam responding ($\beta_{US}$ = `r round(fixef(fit.country_EN_VN)["countryVietnam"], 2)`, p `r ifelse(as.data.frame(summary(fit.country_EN_VN)[["coefficients"]])["countryVietnam", "Pr(>|z|)"] > 0.001, paste0("= ", round(as.data.frame(summary(fit.country_EN_VN)[["coefficients"]])["countryVietnam", "Pr(>|z|)"], 3)), "< .001")`).

@@

For US responding: only the English (EN) corpus is a significant predictor[^3].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^3]: With preregistered exclusion criterion: only the English (EN) and Mandarin (ZH) corpus are significant predictors.

For Vietnam responding: only the Vietnamese (VI) and Mandarin (ZH) corpus are significant predictors[^4].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^4]: With preregistered exclusion criterion: all three corpora are significant predictors.
